

Xieting Zhang, HW2, Biostat 578 Bioinformatics
========================================================

The following codes and annotations include all steps for completing homework 2.

First of all, we need to install bioconductor packages and install the GEOmetadb package which contains the data
we will use. 
```{r}
source("http://bioconductor.org/biocLite.R")
biocLite()
biocLite(c("GEOmetadb", "GEOquery"))
```

Before we can use the GEOmetadb package, we need to call the library
first. 
```{r }
library(GEOmetadb)
```

We also need to make a connection with the database, install limma package and call all the 
libraries we will use. 

```{r }
geo_con <- dbConnect(SQLite(),'GEOmetadb.sqlite')
biocLite("limma")
library(limma)
library(Biobase)
library(data.table)
```


Now we can get the data first:

```{r }
mygds <- getGEO("GSE40812")
```

Since we have downloaded the processed data rather than the raw data, we do not need to normalize it. But we can
clean it up somehow:

```{r }
mygds_new <- mygds
sanitize_pdata <- function(pd){
  keepCols <- c(
    "characteristics_ch1", "characteristics_ch1.2")
  pd <- pd[, keepCols]
  colnames(pd) <- c("pt", "treatment")
  pd$pt <- gsub(".*: ", "", pd$pt)
  pd$treatment<-gsub("treatment: ", "", pd$treatment)
  pd
}
pData(mygds_new[[1]]) <- sanitize_pdata(pData(mygds_new[[1]]))
```

To use the limma package and its regression function, we need to have the model matrix set up with the correct indicators:

```{r }
mm <- model.matrix(~pt+treatment, mygds_new[[1]])
mm[,3] <- mm[,3]+mm[,4]
```

Then we can run the linear fit and use t-test to compute the test scores and compute the p-values for our
coefficient of interest:

```{r }
mfit<- lmFit(mygds_new[[1]], mm[,1:3])
ebay_m <- eBayes(mfit)
topp <- topTable(ebay_m, coef="treatmentPoly IC H", number=Inf, sort.by="none")
```

Now I was stuck and I did not know exactly how to select the 43 probe sets out of the 80 samples I have in 
the data, I could have picked out the adjusted p-values as below:

```{r }
pv<-topp$adj.P.Val
```

But I did not go anywhere with this step of picking out p-values. I think I need some help with this.
Suppose we have been successfull picked out the 43 probe sets that we desire to find, we then can use the 
pheatmap package to plot the expression sets that have p-value < 0.1 as below:

```{r}
install.packages("pheatmap")
library(pheatmap)
## heatmap(exprs(mygds_new[[1]])) This line is not exactly right because we need to pick out the 
## information we need to plot the heatmap, and ignore the other information we do not need.
## exprs() here will extract the expression sets
```