Biostat 578 HW3 Xieting Zhang
========================================================

First of all, get the data from GEO. We have used the data in class we basically do the same as
we did in class: (I called the limma library here as well since we will use it later)

```{r}
source("http://bioconductor.org/biocLite.R")
# install GEO meta database
biocLite("GEOmetadb")
# get GE0 meta database package
library(GEOmetadb)

library(GEOquery)

biocLite("limma")
library(limma)

# Download the mapping information and processed data
gds <- getGEO("GSE29619") 
#main serie #gds[[1]] = LAIV/TIV 0809, gds[[2]] = FACS, gds[[3]] = TIV 0708
```

We need to clean up the data first:

```{r }
### Sanitize data and metadata
gds_new <- gds
sanitize_pdata <- function(pd){
keepCols <- c(
  "characteristics_ch1.1", "characteristics_ch1.2",
  "description", 
  "supplementary_file")
pd <- pd[, keepCols]
colnames(pd) <- c("ptid", "time", "description", "filename")
pd$ptid <- gsub(".*: ", "", pd$ptid)
pd$time <- gsub(".*: ", "", pd$time)
pd$time<-gsub("Day", "D", pd$time)
pd$description<-gsub("(-\\w*){2}$", "", pd$description)
pd$filename<-basename(as.character(pd$filename))
pd$filename<-gsub(".CEL.gz", "", pd$filename)
pd
}

pData(gds_new[[1]]) <- sanitize_pdata(pData(gds_new[[1]]))
pData(gds_new[[2]]) <- sanitize_pdata(pData(gds_new[[2]]))
pData(gds_new[[3]]) <- sanitize_pdata(pData(gds_new[[3]]))
```

Now we can use voom and limma to find genes that are differentially expressed at each time point
compared to baseline:

```{r}
new_set <- ExpressionSet(assayData = exprs(gds_new[[1]])+1) ## add 1 based on the lecture notes
TIV_08 <- new_set
pData(TIV_08) <- pData(gds_new[[1]])


mm_TIV_08 <- model.matrix(~ptid+time, TIV_08) # design matrix
new_set_voom <- voom(TIV_08,design = mm_TIV_08)
lm <- lmFit(new_set_voom, mm_TIV_08)
eb <- eBayes(lm)
```

We can use the topTable() function to find the differentially expressed genes based on
False Discovery Rate less than 0.01:

```{r}
library(data.table)
library(heatmap) ## call this library for later use

topT3 <- data.table(topTable(eb, coef = "timeD3", adjust = 'fdr',p.value = 0.01))

topT7 <- data.table(topTable(eb, coef = "timeD7", adjust = 'fdr',p.value = 0.01))

```

We then take the column corresponding to log fold-change and plot using heatmap:

```{r}
PValue <- matrix(c(topT3$logFC[topT3$adj.P.Val<0.1],topT7$logFC[topT7$adj.P.Val<0.1]),ncol=2)
rownames(PValue) <- rownames(topT3)[topT3$adj.P.Val<0.1]
colnames(PValue) <- c("timeD3","timeD7")

pheatmap(PValue)
```


To do GSEA analysis, we need to install the package first:

```{r}
library(BiocInstaller)
biocLite("GSEABase")
library(GSEABase)
```

Download the gene sets and convert the gene sets to gene indices:

```{r}
c2_set <- getGmt("GSEA-sets/c2.all.v4.0.symbols.gmt")
gene_ids <- geneIds(c2_set)
# Camera requires gene-indices
sets_indices <- symbols2indices(gene_ids, fData(gds_new[[1]])[,"Gene Symbol"])
```

Finding enriched gene sets over time:

```{r}
res <- vector("list",length = 2)
ct <- 1
for(i in c(3,7))
{
  contrast <- paste0("timeD",i)
  cont_matrix <- makeContrasts(contrast, levels=mm_TIV_08)
  res[[ct]] <- camera(new_set_voom, sets_indices, design=mm_TIV_08, contrast=cont_matrix, sort=FALSE)
  ct <- ct+1
}
  
```

The display results using heatmap grouping by timepoint:

```{r}
PValue <- sapply(res, function(x){ifelse(x$Direction=="Up", -10*log10(x$PValue), 10*log10(x$PValue))})
rownames(PValue) <- rownames(res[[1]])
PValue_max <- rowMax(abs(PValue))
PValue_small <- PValue[PValue_max>30, ]
anno <- data.frame(Time=paste0("timeD",c(3,7)))
rownames(anno) <- colnames(PValue_small)  <- paste0("timeD",c(3,7))


pheatmap(PValue_small, cluster_cols=FALSE,show_rownames=FALSE)
```